Almost compatible functions and infinite length games

### 2015-12-03 articles/almost-compatible-functions-and-topological-games/article.tex
clear up A,A' for tactics, marks

  In \cite{clontzMengerGamePreprint} Clontz extended Scheepers' application of
  almost-compatible injections to these game variants as well as
  \(\menGame{\oneptlind\kappa}\).

@@

    %TODO maybe this can be made to work...
    % In \cite{MR1129143} Scheepers used a proof of Todorcevic to demonstrate
    % \(\alcompS\kappa\Rightarrow\pl F \ktactwin2\schFillStrictGame\kappa\).
    % For the convenience of the reader we explicitly demonstrate
    % the generalization as follows.

    % Let \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) witness
    % \(\alcompSP\kappa\), and define \(g_A:A\to\omega\) by
    % \(g_A(\alpha)=f_A(\alpha)-|\{\beta\in A:f_A(\beta)<f_A(\alpha)\}|\).

    % We claim that \(\{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}\)
    % must be finite as it is bounded above by
    % \(\max\{M,f_A(\alpha),f_B(\alpha):f_A(\alpha)\not=f_B(\alpha)\}\)
    % where \(M=f_B(\alpha)\) for some \(\alpha\in B\setminus A\).
    % To see this, let \(f_A(\alpha)=f_B(\alpha)=N>M\) and assume
    % \(f_A(\beta)\not=f_B(\beta)\) implies \(f_A(\beta),f_B(\beta)<N\).
    % Then
    % \[
    %   g_A(\alpha)
    %     =
    %   N-|\{\beta\in A:f_A(\beta)<N\}|
    %     >
    %   N-|\{\beta\in B:f_B(\beta)<N\}|
    %     =
    %   g_B(\alpha)
    % \]
    % with the strictness of the inequality witnessed by \(f_B(\alpha)=M<N\)
    % for some \(\alpha\in B\setminus A\).

    % As a result,
    % \[
    %   \sigma(\<A,B\>)
    %     =
    %   \{\alpha\in A:f_A(\alpha)\not=f_B(\alpha)\}
    %     \cup
    %   \{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}
    % \]
    % is a legal \(2\)-tactic for \(\pl F\). Let \(C=\<C_0,C_1,\dots\>\) be
    % a strictly increasing sequence of countable sets and
    % \(\alpha\in C_n\). Of course if
    % \(f_{C_{n+m}}(\alpha)\not=f_{C_{n+m+1}}(\alpha)\) for some
    % \(m<\omega\) we have \(\alpha\in\sigma(\<C_{n+m},C_{n+m+1}\>)\) immediately.
    % Otherwise let \(f_{C_{n+m}}(\alpha)=f_{C_{n+m+1}}(\alpha)=N\)
    % for all \(m<\omega\).

  It was left open if \(\alcompSP\kappa\) was equivalent to these
  game-theoretic conditions. The answer is consistently negative.

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillIntGame{\omega_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillIntGame{\omega_\alpha}\).
  \end{theorem}

  \begin{proof}
    Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    \(\schFillIntGame{\omega_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    for \(\pl F\) in \(\schFillIntGame{\omega_\alpha}\) as follows:
    \[
      \sigma(\<C\>,0)
        =
      \sigma_0(\<C\cap\omega_{\beta_0}\>,0)
    \]
    \[
      \sigma(\<C,D\>,n+1)
        =
      \sigma_{n+1}(\<D\cap\omega_{\beta_{n+1}}\>,0)
        \cup
      \bigcup_{m\leq n}
      \sigma_m(\<C\cap\omega_{\beta_m},D\cap\omega_{\beta_m}\>,n-m+1)
    \]

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillIntGame{\omega_\alpha}\), and
    \(\alpha\in\bigcap_{n<\omega}C_n\).
    Choose \(N<\omega\) with \(\alpha<\omega_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\omega_{\beta_{N+1}},C_{N+2}\cap\omega_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillIntGame{\omega_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning strategy and
    \(\alpha\in\bigcap_{n<\omega}C_{N+n+1}\cap\omega_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\omega_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\omega_{\beta_{N+1}},C_{N+M+2}\cap\omega_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning strategy.
  \end{proof}

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillInitialGame{\omega_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillInitialGame{\omega_\alpha}\).
  \end{theorem}

@@

  \begin{corollary}
    It is consistent that \(\alcompSP{\aleph_\omega}\) fails,
    but as \(\alcompSP{\aleph_k}\) holds for all \(k<\omega\), we have
    \(\pl F\kmarkwin2\schFillIntGame{\omega_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\omega_\omega}\).
  \end{corollary}

  Our topological question \ref{mainQuestion} remains unsolved; however,
  our results have revealed that we cannot hope to find a \(ZFC\)
  counterexample where \(X=\oneptlind\kappa\). This is
  because if we also assume \(V=L\), it follows that
  \(\alcompSP{\kappa}\) and therefore \(\pl F\kmarkwin2\menGame{X}\)
  for every cardinal.

>>>>

  In \cite{clontzMengerGamePreprint} Clontz extended Scheepers' application of
  almost-compatible injections to these game variants as well as
  \(\menGame{\oneptlind\kappa}\). However, when considering Markov strategies,
  finite-to-one functions suffice.

@@

  We include the following proof from \cite{MR1129143} for the convenience
  of the reader.

  \begin{theorem}
    \(\alcompS\kappa\Rightarrow\pl F \ktactwin2\schFillStrictGame\kappa\)
  \end{theorem}

  \begin{proof}
    Let \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) witness
    \(\alcompS\kappa\), and define \(g_A:A\to\omega\) by
    \(g_A(\alpha)=f_A(\alpha)-|\{\beta\in A:f_A(\beta)<f_A(\alpha)\}|\).

    We claim that \(\{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}\)
    must be finite as it is bounded above by
    \(\max\{M,f_A(\alpha),f_B(\alpha):f_A(\alpha)\not=f_B(\alpha)\}\)
    where \(M=f_B(\alpha)\) for some \(\alpha\in B\setminus A\).
    To see this, let \(f_A(\alpha)=f_B(\alpha)=N>M\) and assume
    \(f_A(\beta)\not=f_B(\beta)\) implies \(f_A(\beta),f_B(\beta)<N\).
    Then
    \[
      g_A(\alpha)
        =
      N-|\{\beta\in A:f_A(\beta)<N\}|
        >
      N-|\{\beta\in B:f_B(\beta)<N\}|
        =
      g_B(\alpha)
    \]
    with the strictness of the inequality witnessed by \(f_B(\alpha)=M<N\)
    for some \(\alpha\in B\setminus A\).

    As a result,
    \[
      \sigma(\<A,B\>)
        =
      \{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}
    \]
    is a legal \(2\)-tactic for \(\pl F\). Let \(C=\<C_0,C_1,\dots\>\) be
    a strictly increasing sequence of countable sets and
    \(\alpha\in C_n\). Noting that \(f_A\) is an injection and not
    just finite-to-one, \(0\leq g_{C_{n+m}}(\alpha)\) for all
    \(m<\omega\), and it follows that
    \(g_{C_{n+m}}(\alpha)\leq g_{C_{n+m+1}}(\alpha)\) for some \(m<\omega\).
    Therefore \(\alpha\in\sigma(\<C_{n+m},C_{n+m+1}\>)\).
  \end{proof}

  So it would seem that \(\alcompSP\kappa\) is sufficient only when
  considering Markov strategies. (Of course,
  \(\alcompSP\kappa\Rightarrow\pl F\kmarkwin2\schFillGame\kappa
  \Rightarrow\pl F\kmarkwin2\schFillStrictGame{\kappa}\).)
  We would like to demonstrate that \(\alcompSP\kappa\) is not
  necessary.

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillIntGame{\aleph_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillIntGame{\aleph_\alpha}\).
  \end{theorem}

  \begin{proof}
    Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    \(\schFillIntGame{\aleph_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    for \(\pl F\) in \(\schFillIntGame{\aleph_\alpha}\) as follows:
    \[
      \sigma(\<C\>,0)
        =
      \sigma_0(\<C\cap\aleph_{\beta_0}\>,0)
    \]
    \[
      \sigma(\<C,D\>,n+1)
        =
      \sigma_{n+1}(\<D\cap\aleph_{\beta_{n+1}}\>,0)
        \cup
      \bigcup_{m\leq n}
      \sigma_m(\<C\cap\aleph_{\beta_m},D\cap\aleph_{\beta_m}\>,n-m+1)
    \]

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillIntGame{\aleph_\alpha}\), and
    \(\alpha\in\bigcap_{n<\omega}C_n\).
    Choose \(N<\omega\) with \(\alpha<\aleph_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\aleph_{\beta_{N+1}},C_{N+2}\cap\aleph_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillIntGame{\aleph_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning strategy and
    \(\alpha\in\bigcap_{n<\omega}C_{N+n+1}\cap\aleph_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\aleph_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\aleph_{\beta_{N+1}},C_{N+M+2}\cap\aleph_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning strategy.
  \end{proof}

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillInitialGame{\aleph_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillInitialGame{\aleph_\alpha}\).
  \end{theorem}

@@

  \begin{corollary}
    It is consistent that \(\alcompSP{\aleph_\omega}\) fails,
    but as \(\alcompSP{\aleph_k}\) holds in \(ZFC\) for all \(k<\omega\),
    both
    \(\pl F\kmarkwin2\schFillIntGame{\aleph_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\aleph_\omega}\) hold in \(ZFC\).
  \end{corollary}

  Note that Question \ref{mainQuestion} remains unsolved; however,
  our results have revealed that we cannot hope to find a \(ZFC\)
  counterexample where \(X=\oneptlind\kappa\). This is
  because if we also assume \(V=L\), it follows that
  \(\alcompSP{\kappa}\) and therefore \(\pl F\kmarkwin2\menGame{X}\)
  for every cardinal. Although, one may consider the following weaker question.

  \begin{question}
    Is there a model of \(ZFC\) where both
    \(\pl F\kmarkwin2\menGame{\oneptlind{\aleph_{\omega+1}}}\) and
    \(\alcompSP{\aleph_\omega}\) fail?
  \end{question}
